---
title: "Exploratory Data Analysis of Blue Water Baltimore's Tree Giveaway data (2013 - 2023)"
author: "Eduardo Marin (with Dexter H Locke)"
date: "`r format(Sys.time())`"
output:
  html_document:
    theme: spacelab
    code_folding: hide
    fig_width: 8
    fig_height: 7
    fig_caption: true
    toc: true
    toc_float: true
    self_contained: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---
### General Outline (with objectives)

**WHEN**

-   Year Map Overview (each color based off of year, plus Shiny app
    (filter for years))\
-   Stacked Bar Chart of Each Year (each season as different colors)

**WHERE**

-   Three Jurisdictions Map Overview (different colors based off of the
    three tree giveaway distribution)\
-   Bar Chart (different colors based on the locations with count)\
-   Baltimore Neighborhood Map Overview (different colors based on
    density of the plantings)\
-   Bar Chart Highlighting Top 10 Planted Neighborhoods (different
    colors based on neighborhood), plus aggregating the rest as Other\
-   Census Tract Overview in Baltimore (different colors based on
    density of the plantings)\
-   Census Tract Overview in Baltimore County (different colors based on
    density of the plantings)\
-   Bar Graph highlighting Top 10 Locations Planted (Cleaned)

**WHAT**

-   Map Overview of Top 10 Species (each species as a different color)\
-   Stacked Bar Chart (Highlighting the different species for each one)\
-   Map Overview of Top 5 Species each Season (each species as a
    different color, RShiny)\
-   Bar Graph of Top 5 Species for Each Year (each species as a color)

**WHO**

-   Map Overview with Planting to Redline Zones (RShiny with Year)\
-   Map Overview with Planting to CBT Zones (RShiny with Year)

```{r message=FALSE, include=FALSE}

# List all of your packages here

knitr::opts_chunk$set(echo = TRUE)

packs <-c(
            'janitor'    # cleans things up, also pipe-friendly cross-tabulations
           , 'sf'         # for spatial data support
          , 'tidyverse'  # cuz
          , 'tidylog'    # prints out what was done in dplyr and tidyr
          , 'magrittr'   # for the pipe
          , 'mapview'    # web maps for zooming and panning around
          #, 'beepr'      # makes noise when things are done!
          , 'tictoc'     # timing things.
          , 'raster'
          # , 'doParallel' # does what is says! PARALLEL
          # 'broom.mixed',# tidiers for mixed models AND nlme::gls()
          # , 'lubridate'   # DATES!
          , 'tidycensus' # tidy census package
          , 'tidygeocoder' # geo coding
          , 'leaflet' #creating the interactive mapping elements (more specific)
          , 'shiny'
          , 'leafsync'  # linked maps
          , 'DT' 
          )     

# IF the packages in 'packs' are not already installed, install them
# OTHERWISE do NOTHING
if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
install.packages(setdiff(packs, rownames(installed.packages())))
}

# lapply(packs, library, character.only = TRUE)
# this actually loads them with library(package_name)
vapply(packs, library, character.only = TRUE, logical(1), logical.return = TRUE, quietly = TRUE)


# # for tidycensus
# setting for get_acs
census_api_key('58fc555c77c229747ade7d9fe50e7c71297cf91a', install = TRUE, overwrite = TRUE)
readRenviron("~/.Renviron")
options(tigris_use_cache = TRUE)
```

```{r include=FALSE}
## B get census data

vars21 <- tidycensus::load_variables('acs5', year = 2021)
# vars21 |> View()

my_vars <- c(  'pop'      = 'B01001_001'
             , 'medhhinc' = 'B19013_001'
             )

(MD <-
  tidycensus::get_acs(
      geography = 'state'
    , state = 'MD'
    , variables = my_vars
    , year = 2021
    , geometry = TRUE
    , output = 'wide'
    , moe_level = 95
    ) |>
   st_transform(crs = st_crs(4326)))

(cbg <-
  tidycensus::get_acs(
      geography = 'block group'
    , state = 'MD'
    , county = c('Baltimore City', 'Baltimore County')
    , variables = my_vars
    , year = 2021
    , geometry = TRUE
    , output = 'wide'
    , moe_level = 95
    ) |>
    st_transform(crs = st_crs(MD)) |>
    mutate(jurisdiction = case_when(
      str_like(GEOID, '24510%') ~ "Baltimore City",
      str_like(GEOID, '24005%') ~ "Baltimore County",
      TRUE ~ "Neither"
    )))

```

```{r message=FALSE, include=FALSE}

# Tree Tribble
species_data <- 
  tribble(
  ~common_name,        ~genus_species,            ~itree,
  "Hackberry",         "Celtis occidentalis",     "CEOC",
  "Willow Oak",        "Quercus phellos",         "QUPH",
  "Black Gum",         "Nyssa sylvatica",         "NYSY",
  "Eastern Redbud",    "Cercis canadensis",       "CECA",
  "Sassafras",         "Sassafras albidum",       "SAAL",
  "Chestnut Oak",      "Quercus prinus",          "QUPR",
  "Eastern Red Cedar", "Juniperus virginiana",    "JUVI",
  "Bald Cypress",      "Taxodium distichum",      "TADI",
  "Black Oak",         "Quercus velutina",        "QUVE",
  "American Holly",    "Ilex opaca",              "ILOP",
  "Sweetbay Magnolia", "Magnolia virginiana",     "MAVI",
  "American Sweetgum", "Liquidambar styraciflua", "LIST",
  "Bur Oak",           "Quercus macrocarpa",      "QUMA1",
  "Black Chokeberry",  "Aronia melanocarpa",      "ARME6",
  "Blackhaw Viburnum", "Viburnum prunifolium",    "VIPR",
  "Gingko",            "Ginkgo biloba",            "GIBI",
  "American Persimmon", "Diospyros virginiana",    "DIVI",
  "Flowering Cherry",   "Prunus yedoensis",        "PRYE",
  "Post Oak",           "Quercus stellata",        "QUST",
  "Chinkapin Oak",      "Quercus muehlenbergii",   "QUMU",
  "Tulip Poplar",       "Liriodendron tulipifera", "LITU",
  "Shortleaf Pine",     "Pinus echinata",          "PIEC",
  "Eastern White Pine", "Pinus americana",         "PIST",
  "Yellow Birch",       "Betula alleghaniensis",   "BEAL2",
  "American Yellowwood","Cladrastis luteau",       "CLLU",
  "Smooth Service Berry","Amelanchier laevis",     "AMLA",
  "Boxelder Maple",    "Acer negundo",             "ACNE",
  "American Linden",   "Tilia americana",          "TIAMRE",
  "American Plum",     "Prunus americana",         "PRAM"
  )
  
#Quick change
quick <- c(1,2,3)

# Loading data
bwb_total_giveaway <-
  read_csv("output_data/bwb_giveaway_total_final_2024-05-13.csv") |>
  filter(!is.na(lat)) |> #Good double check
  st_as_sf(coords = c("long","lat"), crs = 4326) |> 
  st_join(cbg) |> 
  mutate(common_name = case_when(
    common_name == "Boxelder" ~ "Boxelder Maple",
    common_name == "Yoshino Flowering Cherry" ~ "Flowering Cherry",
    common_name == "Black Tupelo" ~ "Black Gum",
    common_name == "Chinquapin" ~ "Chinkapin Oak",
    common_name == "Breakfast" ~ "Other",
    common_name == "Blackhaw Vibernum" ~ "Blackhaw Viburnum",
    common_name == "Ginkgo" ~ "Gingko",
    TRUE ~ common_name
  )) |>
  filter(address != "822 Montpellier") |> 
  left_join(species_data, by = "common_name") |>
  mutate(
    genus_species = coalesce(genus_species.x, genus_species.y),
    itree = coalesce(itree.x, itree.y)
  ) |>
  dplyr::select(-genus_species.x, -genus_species.y, -itree.x, -itree.y) |>
  left_join(species_data, by = "genus_species") |>
  mutate(
    common_name = coalesce(common_name.x, common_name.y),
    itree = coalesce(itree.x, itree.y)
  ) |>
  dplyr::select(-common_name.x, -common_name.y, -itree.x, -itree.y) |>
 mutate(genus_species = ifelse(common_name == "Other", "Other", genus_species),
       itree = ifelse(common_name == "Other", NA_character_, itree))
         
```

# 1 When

```{r eval=FALSE, include=FALSE}
# Defining the UI for Shiny component
year_ui <- fluidPage(
  titlePanel("BWB Tree Giveaway Interactive Map with Year Filter"),
  sidebarLayout(
    sidebarPanel(
      selectInput("year", "Select Year:", choices = unique(bwb_total_giveaway$year))
    ),
    mainPanel(
      leafletOutput("interactive_map")
    )
  )
)

# Creating the color palette of the webmap

# Creating the webmap with the Shiny component
server <- function(input, output, session) {
  filtered_data <- reactive({
    bwb_total_giveaway |> filter(year == input$year)
  })
  
  output$interactive_map <- renderLeaflet({
    leaflet() |>
      addProviderTiles(providers$CartoDB.Positron) |>
      addCircleMarkers(
        data = filtered_data(),
        radius = 2,
        fillOpacity = 0.1,
        color = "#2E6F40",
        popup = ~paste(
          "<b>Year:</b>", year, "<br>",
          "<b>Season:</b>", season, "<br>",
          "<b>Date:</b>", date, "<br>",
          "<b>Common Name:</b>", common_name, "<br>",
          "<b>Scientific Name:</b>", genus_species, "<br>",
          "<b>Address:</b>", address, "<br>",
          "<b>City:</b>", city, "<br>",
          "<b>State:</b>", state, "<br>",
          "<b>Zip Code:</b>", zip_code, "<br>"
        )
      )
  })
}

# Run the application
shinyApp(year_ui, server)


# General map of all points
general_map <- 
  leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) #the basemap

# Customizing the general map
general_map <- 
  general_map |>
  addCircleMarkers(
  data = bwb_total_giveaway,
  radius = 2, 
  fillOpacity = 0.1,
  color = "#2E6F40",
  popup = ~paste(
    "<b>Year:</b>", year, "<br>",
    "<b>Season:</b>", season, "<br>",
    "<b>Date:</b>", date, "<br>",
    "<b>Common Name:</b>", common_name, "<br>",
    "<b>Scientific Name:</b>", genus_species, "<br>",
    "<b>Address:</b>", address, "<br>",
    "<b>City:</b>", city, "<br>",
    "<b>State:</b>", state, "<br>",
    "<b>Zip Code:</b>", zip_code, "<br>"
  )
    )

# Displaying the map
general_map
```

## A General Map

# 3 Performing the a-spatial double checks

## A Time

```{r}
# Playing with year (result is season wide)
bwb_total_giveaway |> 
  st_drop_geometry() |>
  tabyl(year, season) |>
  as_tibble() |>
  DT::datatable()
  
# Same data, different shape (result is time long)
(giveaway_counts_season_year <-
  bwb_total_giveaway |> 
  st_drop_geometry() |>
  group_by(year, season) |>
  count())

# # Graphing this giveaway counts
# giveaway_counts_season_year |>
#   ggplot(aes(x = year, y = n, color = season)) +
#   geom_point() +
#   # geom_line() +
#   geom_smooth(method = "lm") +
#   geom_smooth()

#geom_sf is for vector spatial data, data is the first argument, matters when you have layers. In terms of the ggplot, aes defines the conditions of the graph (let's you pass it through each point). geom_point makes the point graph. ggplot can be seen as global 

#NULL Lets you always rull

# Graph about time
giveaway_counts_season_year |>
  ggplot(aes(x = as.factor(year), y = n, color = season, group = season)) +
  geom_point() +
  geom_line() +
  labs(
      title = "Number of Tree Giveaways by Year"
    , subtitle = "The Blue Water Baltimore data is also looking at specific seasons"
    , caption = "Testing"
    , x = "Year"
    , y = "Number of Trees Given Away"
  ) +
  theme_bw(16) +
  theme(legend.position = c(.1, .7)) +
  NULL
  

```

## B Location

```{r}

glimpse(bwb_total_giveaway)

(location_counts <- 
  bwb_total_giveaway |>
  st_drop_geometry() |>
  tabyl(giveaway_location) |> 
  as_tibble() |> 
  arrange(desc(n))
)

location_counts |> 
  ggplot(aes(n, reorder(giveaway_location, n))) + # pushes the bars in the desc order
  geom_col() +
  theme_bw(16) +
  labs(y = 'Giveaway Location'
       , x = 'Number of trees') +
  NULL

```

## C Species

```{r}
# 2x check
bwb_total_giveaway |>
  filter(is.na(common_name) | is.na(genus_species))

#Making that
bwb_total_giveaway |>
  filter(is.na(common_name) & is.na(genus_species))

#Genus Species
bwb_total_giveaway |>
  st_drop_geometry() |>
  tabyl(genus_species) 

#Common Name
(giveaway_counts_common_name <-
  bwb_total_giveaway |>
  st_drop_geometry() |>
  tabyl(common_name) |>
  as_tibble() |>
  arrange(desc(n)) |> 
  mutate(group = rep(c("common", "rare"), each = 42))
)


giveaway_counts_common_name |> 
  ggplot(aes(n, reorder(common_name, n))) + # pushes the bars in the desc order
  geom_col() +
  theme_bw(10) +
  labs(y = 'Common Name'
       , x = 'Number of trees'
       , title = 'Which tree species were most and least common?'
       , subtitle = paste0('lots of Serviceberry... they make up '
                           , round(giveaway_counts_common_name |> slice(1) |> pull(percent) * 100, 2)
                           , ' percent of all giveaways')
       ) +
  facet_wrap(~group, scales = 'free') +
  NULL

```

## D Making the count by city

```{r}
(giveaway_counts_city_season <-
  bwb_total_giveaway |>
  st_drop_geometry() |>
  group_by(city, year, season) |>
  count() |>
  as_tibble() |>
  arrange(desc(n))
)

giveaway_counts_city_season |>
  ggplot(aes(year, n, color = season, group = season)) +
  geom_point() +
  geom_line() +
  facet_wrap(~city) +
  theme_bw(16) +
  NULL

## 
  bwb_total_giveaway |>
  st_drop_geometry() |>
  group_by(zip_code, year, season) |>
  count() |>
  as_tibble() |>
  arrange(desc(n)) |>
  ggplot(aes(year, n, color = season, group = season)) +
  geom_point() +
  geom_line() +
  facet_wrap(~zip_code) +
  theme_bw(16) +
  NULL

## Comparing to county data
  bwb_total_giveaway |>
  st_drop_geometry() |>
  group_by(county, year, season) |>
  count() |>
  as_tibble() |>
  arrange(desc(n)) |>
  ggplot(aes(year, n, color = season, group = season)) +
  geom_point() +
  geom_line() +
  facet_wrap(~county) +
  theme_bw(16) +
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) +
  NULL
  
```

# 4 maps and graphs to make

```{r, fig.height=4}

# Species over time
# graph with time on horizontal
# vertical is amount
# color for season
# lines per species (too many)
bwb_total_giveaway |> 
  st_drop_geometry() |> 
  group_by(year, season, common_name) |> 
  count() |> # summary() # n ranges from 1 to 192
  filter(n > 10) |> 
  ggplot(aes(year, n, group = common_name, color = season)) +
  geom_point() +
  geom_line() +
  facet_wrap(~common_name)


# map for the top five species
five_most_common_species <- 
  bwb_total_giveaway |> 
  st_drop_geometry() |> 
  group_by(common_name) |> 
  count() |> 
  ungroup() |> 
  arrange(desc(n)) |> 
  slice(1:10)

bwb_total_giveaway |> 
  filter(common_name %in% five_most_common_species$common_name) |> 
  mapview()


bwb_total_giveaway |> 
  filter(common_name %in% five_most_common_species$common_name) |> 
  mapview(zcol = 'common_name')


bwb_total_giveaway |> 
  filter(common_name %in% five_most_common_species$common_name) |> 
  mapview(zcol = 'common_name', layer.name = "Common Name")


bwb_total_giveaway |> 
  filter(common_name == 'Serviceberry') |> 
  mapview(zcol = 'year', layer.name = "Common Name")

```

# old sandbox

\`\`\`{r eval=FALSE, message=FALSE, include=FALSE} \# 5 augment tree
data with census

bwb_total_giveaway_aug \<- bwb_total_giveaway \|\> st_join(cbg \|\>
tidylog::select(-NAME) )

bwb_total_giveaway_aug \|\> glimpse()

bwb_total_giveaway_aug \|\> mapview(zcol = 'medhhincE')

# Create an interactive map

bwb_total_giveaway \|\> mapview()

# Messing with data

bwb_total_giveaway \|\> filter(year == 2015) \|\> mapview()

# Messing with data

bwb_total_giveaway \|\> mapview(zcol = "year")

# Display the map

map_giveaway

#Final squeaky cleaned geocoded and get the GItHub resolved, and by the
wenday. Email how you going. Sit next. Finish it by next Monday:
Geocoding and GitHUb Meet by Zoom Tuesdya from 9 to 11
